---
title: "Analysis"
author: "Jiří Klepl"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

## Load libraries

```{r load libraries, warning = F, message = F}
library(raster)
library(imager)
library(tidyverse)
library(here)
library(ez)
library(saccades)
require(ggplot2) 

library(papaja)


source("solution.R") # functions are in separate file for easier debugging
```

## Download data

Download zip file from the web page. Description of the dataset is on the page [ETRA dataset description](https://etra.acm.org/2019/challenge.html)

```{r download data}
data_pth <- "data"
if(!dir.exists(data_pth)) {
  dir.create(data_pth)
}
# uncomment following file for downloading data
# download.file("http://smc.neuralcorrelate.com/ETRA2019/ETRA2019Challenge.zip",here(data_pth,"ETRA2019Challenge.zip"))
```

Than unzip data and move them to the folder. The directory *data* should therefore contain following directories/files

- data
- images
- DataSummary.csv

## Load data

You can use *readr* package, which is supplied within *tidyverse* package bundle. Multiple files can be read using these lines of code. For example, we want only load Free viewing condition in Puzzle subtask for subject 9

```{r load data, warning=F, message = F}
# local_data_pth <- here(data_pth, "data", "009")
# files <- dir(local_data_pth, pattern = ".*Fixation_Puzzle_.*.csv") # get file names

# df_fix_puzzle <- files %>%
#   map(read_data) %>% 
#   reduce(rbind) %>% as_tibble()

# local_data_pth <- here(data_pth, "data", "009")
# files <- dir(local_data_pth, pattern = ".*Fixation_Waldo_.*.csv") # get file names

# df_fix_waldo <- files %>%
#   map(read_data) %>% 
#   reduce(rbind) %>% as_tibble()

```

You can easily adjust this code for your hypotheses.

## Hypotheses

All the hypotheses apply to the "puzzle task" experiments:

1. Is the length of fixations dependent on whether it is directed at a location close to the differences in the puzzle experiments?
2. Is the length of fixations dependent on the distance from the boundary of the image
3. Is there any significant dependence of pupil dilatation on time

## Data manipulation

In the whole analysis for all hypotheses we simplify the data by taking averages of values of both eyes instead of the values themselves wherever applicable (pupil dilatations, x/y locations, etc.).

In the analysis for the first two hypotheses we use the provided method for detecting fixations, for the first hypothesis we compute the difference of the left and the right image, this is done by taking the top 3% percentile of pixels with the most color distance (using euclidean metric). This leaves a lot of edges and textured areas as false positives so the script uses "CannyEdge" edge detection to partially correct this.

But even with the false positives we should get at least partially consistent correlation if the hypotheses are true.

```{r data manipulation, warning = F, message = F}
first_hyp()
second_hyp()
third_hyp()
third_hyp_two()
```

## Results

The analysis for the first two hypotheses shows no apparent correlation so there is no further research on them.

The analysis for the third hypothesis shows very strong correlation between time and pupil dilatation; and the t.test applied on the collected data (for each experiment, then for all each participant's experiments and finally for all experiments) confirms that the correlation isn't due to randomness (with probability consistently below reasonable threshold).

applying `lm` to the data collected by the third hypothesis gives us an approximation of the relation between the values as follows: `pupil dilatation = 1352.27 + .0054 * time`.

So we can conclude that the length of fixations isn't dependent on the distance from the target locations nor on the distance from the boundaries of the stimulus images; and that the pupil dilatation is indeed dependent on time.
